---
title: "Workout1"
output:  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
library(dplyr)
library(ggplot2)
library(readr)
```


### Data Preparation
```{r}



nba2018<-read_csv('../data/nba2018.csv',
                  col_types= cols(
                    .default = col_integer(),
                    player = col_character(),
                    number = col_character(),
                    team = col_character(),
                    position = col_factor(c("C","PF","PG","SF","SG")),
                    height = col_character(),
                    birth_date = col_character(),
                    country = col_character(),
                    experience = col_character(),
                    college = col_character(),
                    salary = col_double(),
                    field_goals_perc = col_double(),
                    points3_perc = col_double(),
                    points2_perc = col_double(),
                    effective_field_goal_perc = col_double(),
                    points1_perc = col_double()
                  ))
```
## experience:
```{r}  
factor(nba2018$experience)
nba2018$experience[nba2018$experience == "R"] <- 0
as.integer(nba2018$experience)
```
## salary:
```{r}  
nba2018$salary<- nba2018$salary/1000000
```
## Position:
```{r}
nba2018$position<-factor(nba2018$position, labels = c("center","power_fwd","point_guard","small_fwd",'shoot_guard'))
```
```{r}  
nba2018new<- mutate(nba2018,
                    missed_fg = points3_atts - points3 + points2_atts - points2,
                    missed_ft = points1_atts - points1,
                    rebounds = off_rebounds + def_rebounds,
                    efficiency = (points + rebounds + assists + steals + blocks
                                  - missed_fg - missed_ft - turnovers)/ games )
```
## sink
```{r}
sink('../output/efficiency-summary.txt')
summary(nba2018new$efficiency)
```
## creating nba2018.csv
```{r}
teamstats<-summarise( group_by(nba2018new, team),
                      experience = sum(as.integer(experience)),
                      salary= sum(salary),
                      points3byteam =  sum(points3),
                      points2byteam = sum(points2),
                      freebyteam =  sum(points1),
                      pointsbyteam =  sum(points),
                      offrebounds= sum(off_rebounds),
                      defrebounds= sum(def_rebounds),
                      assistteam = sum(assists),
                      stealsteam = sum(steals), 
                      blocksteam=  sum( blocks),
                      turnovers = sum(turnovers),
                      foulsteam = sum(fouls),
                      efficiency = sum(efficiency))
```
## sink
```{r}
sink('../data/teams-summary.txt')
summary(teamstats)
```
```{r}
write_csv(teamstats, '../data/nba2018-teams.csv')
```



```{r}
teamsalary<-teamstats %>% 
  arrange(salary)

teamsalary$num<- 1:nrow(teamstats)
```

```{r}
ggplot(data= teamsalary, 
       aes(x= reorder(team,num),
           y = salary)
       )+
  geom_bar(stat = 'identity')+
  coord_flip()+
  geom_hline(yintercept= mean(teamsalary$salary), col= 'red')
  
```
```{r}
teamspoints<-teamstats %>% 
  arrange(pointsbyteam)

teamspoints$num<- 1:nrow(teamstats)
```
```{r}
ggplot(data= teamspoints, 
       aes(x= reorder(team,num),
           y = pointsbyteam)
       )+
  geom_bar(stat = 'identity')+
  coord_flip()+
  geom_hline(yintercept= mean(teamspoints$pointsbyteam), col= 'red')
```
```{r}
teamefficiency<-teamstats %>% 
  arrange(efficiency)

teamefficiency$num<- 1:nrow(teamstats)
```
```{r}
ggplot(data= teamefficiency, 
       aes(x= reorder(team,num),
           y = efficiency)
       )+
  geom_bar(stat = 'identity')+
  coord_flip()+
  geom_hline(yintercept= mean(teamefficiency$efficiency), col= 'red')
```


I will use assists per turnover to rank a team
which means the more assists per turnover by a team, the better coordinated that team is.


```{r}
teamstatsnew<-mutate(teamstats, APT = assistteam/turnovers)
teamassistoverturn<-teamstatsnew %>% 
  arrange(APT)

teamassistoverturn$num<- 1:nrow(teamstatsnew)
```
```{r}
ggplot(data= teamassistoverturn, 
       aes(x= reorder(team,num),
           y = APT)
       )+
  geom_bar(stat = 'identity')+
  coord_flip()+
  geom_hline(yintercept= mean(teamassistoverturn$APT), col= 'red')
```

### comments and reflections

this was the first time working on a project with such file structures
its really complicated
Not the first time using relative path
yes first time r script
things were sooo hard
everything was pretty hard except for the repetitive things
No help
spent 8 hours doing this
except for the ggplot nothing was interesting